---
title: "**treeheatr**: an introduction"
output: rmarkdown::github_document
---

[![vignette](https://img.shields.io/badge/-Vignette-green?logo=spinnaker)](https://trang1618.github.io/treeheatr/articles/explore.html)
[![documentation](https://img.shields.io/badge/-Documentation-purple?logo=read-the-docs)](https://trang1618.github.io/treeheatr/reference/)
[![github-action-status](https://img.shields.io/github/workflow/status/trang1618/treeheatr/R-CMD-check?logo=github)](https://github.com/trang1618/treeheatr/actions)
`r badger::badge_cran_download("treeheatr", "grand-total", "blue")`
`r badger::badge_doi("10.1093/bioinformatics/btaa662", "yellow")`

```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.path = "man/figures/"
)
```

#### *Your decision tree may be cool, but what if I tell you you can make it hot?*

## Changes in treeheatr 0.2.0

The first argument of `heat_tree()`, `data` is now replaced with `x`, 
which can be a dataframe (or tibble), 
a party (or constparty) object specifying the precomputed tree,
or partynode object specifying the customized tree. 
`custom_tree` argument is no longer needed.

## Install 

Please make sure your version of R >= 3.5.0 before installation.

You can install the released version of **treeheatr** from CRAN with:
```{r, eval=FALSE}
install.packages('treeheatr')
```

Or the development version from GitHub with remotes:
```{r, eval=FALSE}
# install.packages('remotes') # uncomment to install devtools
remotes::install_github('trang1618/treeheatr')
```

## Examples

### Penguin dataset

These data were collected and made available by [Dr. Kristen Gorman](https://www.uaf.edu/cfos/people/faculty/detail/kristen-gorman.php) and the [Palmer Station, Antarctica LTER](https://pal.lternet.edu/).

Classification of different types of penguin species.

```{r, fig.height=3.5, message=FALSE, warning=FALSE, dpi = 200}
library(treeheatr)

heat_tree(penguins, target_lab = 'species')
```

### Wine recognition dataset

Classification of different cultivars of wine.

```{r fig.height=3.5, dpi = 200}
heat_tree(wine, target_lab = 'Type', target_lab_disp = 'Cultivar')
```

## Citing treeheatr

If you use treeheatr in a scientific publication, please consider citing the following paper:

Le TT, Moore JH. 
[treeheatr: an R package for interpretable decision tree visualizations](https://doi.org/10.1093/bioinformatics/btaa662). 
_Bioinformatics_. 2020 Jan 1.

BibTeX entry:
```bibtex
@article{le2020treeheatr,
  title={treeheatr: an R package for interpretable decision tree visualizations},
  author={Le, Trang T and Moore, Jason H},
  journal={Bioinformatics},
  year={2020},
  doi="10.1093/bioinformatics/btaa662"
}
```

## How to Use

**treeheatr** incorporates a heatmap at the terminal node of your decision tree.
The basic building blocks to a **treeheatr** plot are (yes, you guessed it!) a decision tree and a heatmap.

* The decision tree is computed with `partykit::ctree()` and plotted with the well-documented and flexible [**ggparty**](https://cran.r-project.org/package=ggparty/) package.
The tree parameters can be passed to **ggparty** functions via the `heat_tree()` and `draw_tree()` functions of **treeheatr**.
More details on different **ggparty** *geoms* can be found [here](https://github.com/martin-borkovec/ggparty).

* The heatmap is shown with `ggplot2::geom_tile()`.
The user may choose to cluster the samples within each leaf node or the features across all samples.

Make sure to check out the [vignette](https://trang1618.github.io/treeheatr/articles/explore.html) for detailed information on the usage of **treeheatr**.

Please [open an issue](https://github.com/trang1618/treeheatr/issues/new) for questions related to **treeheatr** usage, bug reports or general inquiries.

Thank you very much for your support!
