---
title: "**treeheatr**: an introduction"
output: rmarkdown::github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.path = "man/figures/"
)
```

#### *Your decision tree may be cool, but what if I tell you you can make it hot?*

## Install 

Please make sure your version of R >= 3.5.0 before installation.

```{r, eval=FALSE}
devtools::install_github('trang1618/treeheatr')
```

## Examples

### Iris dataset

Classification of different types of iris plants.

```{r, fig.height=3.5, message=FALSE, warnings = FALSE}
library(treeheatr)

grid.draw(heat_tree(iris, class_lab = 'Species'))
```

### Wine recognition dataset

Classification of different cultivars of wine.

```{r fig.height=3.5}
grid.draw(heat_tree(wine, class_lab = 'Type'))
```


## How to Use

**treeheatr** incorporates a heatmap at the terminal node of your decision tree.
The basic building blocks to a **treeheatr** plot are (yes, you guessed it!) a decision tree and a heatmap.

* The decision tree is computed with `partkit::ctree()` and plotted with the well-documented and flexible [**ggparty**](https://cran.r-project.org/web/packages/ggparty/index.html) package.
The tree parameters can be passed to **ggparty** functions via the `heat_tree()` and `draw_tree()` functions of **treeheatr**.
More details on different **ggparty** *geoms* can be found [here](https://github.com/martin-borkovec/ggparty).

* The heatmap is shown with `ggplot2::geom_tile()`.
The user may choose to cluster the samples within each leaf node or the features across all samples.

Make sure to check out the [vignette](https://trang1618.github.io/treeheatr/articles/explore.html) for detailed information on the usage of **treeheatr**.

Please [open an issue](https://github.com/trang1618/treeheatr/issues/new) for questions related to **treeheatr** usage, bug reports or general inquiries.

Thank you very much for your support!

