---
title: "treeheatr"
output: rmarkdown::github_document
---

Integrating heatmap in decision tree.

## Install 

Please make sure your version of R >= 3.6.0 before installation.

```{r, eval=FALSE}
devtools::install_github('trang1618/treeheatr',
                         dependencies = TRUE)
```

## Examples

### Iris dataset

Classification of different types of iris plants.

```{r, fig.height=3.5, message=FALSE, warnings = FALSE}
library(treeheatr)

heat_tree(iris, class_lab = 'Species') %>% 
  grid.draw()
```

### Wine recognition dataset

Classification of different cultivars of wine.

```{r fig.height=3.5}
data(wine, package = 'rattle')

heat_tree(wine, class_lab = 'Type') %>% 
  grid.draw()
```


## How to Use

**treeheatr** incorporates a heatmap at the terminal node of your decision tree.
The basic building blocks to a **treeheatr** plot are (yes, you guessed it!) a decision tree and a heatmap.

* The decision tree is computed with `partkit::ctree()` and plotted with the well-documented and flexible [**ggparty**](https://cran.r-project.org/web/packages/ggparty/index.html) package.
The tree parameters can be passed to **ggparty** functions via the `heat_tree()` and `draw_tree()` functions of **treeheatr**.
More details on different **ggparty** *geoms* can be found [here](https://github.com/martin-borkovec/ggparty).

* The heatmap is shown with `ggplot2::geom_tile()`.
The user may choose to cluster the samples within each leaf node or the features across all samples.

Make sure to check out the [wiki](https://github.com/trang1618/treeheatr/wiki) or the [vignettes](https://github.com/trang1618/treeheatr/vignettes) for detailed information on the usage of **treeheatr**.  

Please [open an issue](https://github.com/trang1618/treeheatr/issues/new) for questions related to **treeheatr** usage, bug reports or general inquiries.
Thank you very much for your support!

