---
title: "treeheatr"
output: rmarkdown::github_document
---

Integrating heatmap in decision tree.

## Install 
```{r, eval=FALSE}
devtools::install_github('trang1618/treeheatr',
                         dependencies = TRUE)
```

## Example
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      dpi = 300,
                      fig.path = "man/figures/README-")
options(warn = -1)
```


```{r, warning = FALSE, message = FALSE, dpi = 700}
library(treeheatr)
library(gtable)

dat_raw <- iris

my_fig <- heat_tree(dat_raw, class_lab = 'Species')

grid::grid.draw(my_fig)
```
  
## How to Use

**treeheatr** incorporates a heatmap at the terminal node of your decision tree.
The basic building blocks to a **treeheatr** plot are (yes, you guessed it!) a decision tree and a heatmap.

* The decision tree is computed with `partkit::ctree()` and plotted with the very well-documented and flexible **ggparty** package.
The tree parameters can be passed to **ggparty** functions via the `heat_tree()` and `draw_tree()` functions of **treeheatr**.
More details on different **ggparty** *geoms* can be found [here](https://github.com/martin-borkovec/ggparty).

* The heatmap is shown with `ggplot2::geom_tile()`.
The user may choose to cluster the samples within each leaf node or the features across all samples.

Make sure to check out the [**wiki**](https://github.com/trang1618/treeheatr/wiki) or the vignettes for detailed information on the usage of **treeheatr**.  

Please [open an issue](https://github.com/trang1618/treeheatr/issues/new) for questions related to **treeheatr** usage, bug reports or general inquiries.
Thank you very much for your support!

