---
title: "Explore **treeheatr**"
author: "Trang Le"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Explore treeheatr}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[UTF-8]{inputenc}
  %\VignetteDepends{cowplot}
---

**treeheatr** displays a more interpretable decision tree visualization by integrating a heatmap at its terminal nodes.
Let's explore the package **treeheatr** a little deeper and see what it can do!


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(treeheatr)
library(dplyr)
library(cowplot)
```

Let's start with the iris dataset!
Running the `heat_tree()` function can be as simple as:

```{r fig.height=3, fig.width=5}
dat_raw <- iris
heat_tree(dat_raw, class_lab = 'Species') %>% 
  grid::grid.draw()
```

But we can adjust a few graphical parameters and add a custom layout for a subset of the nodes by specifying it in the `custom_layout` parameter.
We can also relax the p value threshold to include more features that might be less important in classifying the samples.

```{r fig.height=3, fig.width=5}
heat_tree(
  dat_raw, class_lab = 'Species', 
  custom_layout = data.frame(id = 1, x = 0.8, y = 1),
  p_thres = 1,
  panel_space = 0.01, class_space = 0.2, tree_space_bottom = 0.1, heat_rel_height = 0.4) %>% 
  grid::grid.draw()
```



We can also customize our heattree by passing parameters through to different **ggparty** geoms.
These list parameters are named `*_vars`.
For example:

```{r fig.height=4, fig.width=7}
heat_tree(
  dat_raw, class_lab = 'Species',
  par_node_vars = list(
    label.size = 0.2,
    label.padding = unit(0.1, "lines"),
    line_list = list(
      aes(label = paste("Node", id)),
      aes(label = splitvar),
      aes(label = paste("p =", formatC(p.value, format = "e", digits = 2)))),
    line_gpar = list(
      list(size = 8),
      list(size = 8),
      list(size = 6)),
    id = 'inner'),
  terminal_vars = list(size = 0),
  edge_vars = list(size = 1, color = 'grey')) %>% 
  grid::grid.draw()
```

## Some more examples

```{r fig.height=4, fig.width=7}
heat_tree(wine, class_lab = 'Type') %>% 
  grid::grid.draw()
```

```{r warning=F, fig.height=7, fig.width=7}
p <- lapply(list(flags, saheart, diabetes), heat_tree, class_lab = 'target')

cowplot::plot_grid(plotlist = p, ncol = 1)
```


### Extreme examples

These visualizations may not be very interpretable but serves the purpose of showing the default node layout when the tree grows in size.
The implemented smart layout weighs the x-position of the parent node according to the level of the child nodes as to avoid crossing of tree branches.
This relative weight can be adjusted with the `lev_fac` parameter in `heat_tree()`.
The default `lev_fac = 1.3` seems to provide aesthetically pleasing trees, independent of the tree size.

```{r fig.height=6, fig.width=9}
# waveform has 5000 observations and 40 features
# so this computation may take up to a minute or two.

heat_tree(waveform, class_lab = 'target', heat_rel_height = 0.35) %>%
  grid::grid.draw()
```

```{r fig.height=6, fig.width=9}
heat_tree(wine_quality_red, class_lab = 'target') %>% 
  grid::grid.draw()
```

`lev_fac = 1` draws parent node perfectly in the middle of child nodes (note a few branch crossing).

```{r fig.height=6, fig.width=9}
heat_tree(wine_quality_red, class_lab = 'target', lev_fac = 1) %>% 
  grid::grid.draw()
```


## Mixed data types and other features

**treeheatr** supports mixed data types.

For continuous variables, we can choose to either *scale* (subtract the mean and divide by the standard deviation) or *normalize* (subtract the min and divide by the max) each variable.
Depending on what we want to show in the heatmap, one transformation method can be more effective than the other.
Details on the strengths and weaknesses of different types of data transformation for heatmap display can be found in [this vignette]( https://cran.r-project.org/web/packages/heatmaply/vignettes/heatmaply.html#data-transformation-scaling-normalize-and-percentize) of the **heatmaply** package.

Clustering

Let's take a closer look at how it handles a dataset like Titanic with categorical variables (`Class`, `Sex`, `Embarked`) and continuous variables (others).


```{r fig.height=3.5, fig.width=7}
dat_raw <- titanic %>%
  dplyr::select(Survived, Pclass, Sex, Age, SibSp, Parch, Fare, Embarked) 

heat_tree(
  dat_raw,
  class_lab = 'Survived',
  label_map = c(`1` = 'Survived', `0` = 'Deceased')) %>% 
  grid::grid.draw()
```

